---
title: Cleaning Contact Info
author: ~
date: '2016-11-02'
slug: ''
categories: []
tags:
  - r-code
  - data-cleaning
  - crm
description: ''
draft: true
output:
  blogdown::html_page:
    toc: true
---

```{r setup, include = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  cache = TRUE,
  fig.align = "center"
)
```

## Where are the Examples?

R is great for cleaning data and that's why I was surprised that I couldn't find 
very many examples or even a full-fledged package for cleaning contact information 
(phone numbers, addresses, etc.). Instead I decided to write a few functions of 
my own to get the job done. I wanted to put them front and center as a table 
of contents at the top for people who want to quickly reference them on this page. 
At some point I plan on consolidating CRM-like R code into a package that can 
help others who are working with customer data to quickly clean and analyze using R. 

## Cleaning Functions Framework

When I set out to write cleaning functions I wanted to write them in a way that 
they could be chained together as part of `dplyr` pipeline so that readers of 
my code could see what steps I took to process the data. For this reason I created 
functions that accept a `data.frame` and a target column by name to operate on and 
then return the cleaned `data.frame`. The flow could be something like: 

```{r cleaning-flow, eval=FALSE}
clean_data <- data %>%
  clean_phone_number(targetColumn='Phone') %>%
  clean_phone_number(targetColumn='CellPhone') %>%
  clean_url(targetColumn='URL') %>%
  clean_zip_code(targetColumn='Zip') %>%
  clean_email(targetColumn='Email')
```

Below are functions that I've created to perform the types of cleaning demonstrated 
above.

## The Functions

### Cleaning Phone Numbers

```{r cleaning-phone-numbers}
2 + 2
```

### Cleaning URLs

```{r cleaning-urls}
2 + 2
```

### Cleaning Zip Codes

```{r cleaning-zip-codes}
2 + 2
```

### Cleaning Email Addresses

```{r cleaning-email-addresses}
2 + 2
```
