---
title: "PFA with the R Package aurelius"
author: "Steven M. Mortimer"
date: '2017-02-13'
aliases: 
  - /2017/02/13/pfa-with-the-r-package-aurelius
description: 'How to generate R models into the Portable Format for Analytic (PFA)'
slug: pfa-with-the-r-package-aurelius
categories:
  - R
tags:
- r-code
- PFA
- machine-learning
output:
  blogdown::html_page
---

```{r setup, include = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  cache = TRUE,
  fig.align = "center"
)
```

I first heard about the Portable Format for Analytics (PFA) through KDnuggets in 
2016. There was a <a href="http://www.kdnuggets.com/2016/01/portable-format-analytics-models-production.html" target="_blank">post</a> that in vague way talked about the virtues and benefits of PFA, but not to completely 
supplant PMML, which I was already familiar with. PMML was pretty cool, but I 
couldn't find good support and felt that the XML-based approach was clumsy and 
outdated so when I heard about PFA, I was pretty excited. A new JSON-based standard 
for encoding any type of analytics, not just certain supported models like PMML. 
I mainly work in the R programming language, so I was disappointed to find resources 
lacking that would convert models made in R into the PFA standard. The great 
folks at the <a href="http://dmg.org" target="_blank">Data Mining Group</a> (DMG) had started on an R package 
called `aurelius` that contained the bare bones functions to construct the syntax. 
After reading through the Project pages on GitHub, I saw everything was there, 
but it wasn't as convenient as I would like so I started writing "model producers". 
Model producers are functions that convert model objects in R into their PFA 
equivalent to be exported and used in other systems. In this process I learned 
a great deal. I thought I knew how models worked, but actually reading through 
the mechanics of scoring has made me appreciate them so much more. Below is an 
outline of working with the `aurelius` package, but note that this work has yet 
to be integrated into the project and exists on forked branch in my GitHub 
(installation command shown below).

```{r install-package, eval=FALSE}
# you only need to run this once because it 
# installs the package
devtools::install_github('StevenMMortimer/hadrian', 
                         ref='feature/add-r-package-structure', 
                         subdir='aurelius', 
                         quick=T) 
```

In this example we'll load the randomForest package to demonstrate how to convert 
randomForest models into PFA. Currently, the package converts models of many types: 
`lm`, `glm`, `glmnet`, `gbm`, `randomForest`, `naivebayes`, `lda`, `kcca`, and `kmeans`.

```{r load-packages, warning=FALSE, message=FALSE}
options(stringsAsFactors = FALSE)
library(aurelius)
library(randomForest) 
```

## PFA Structure

Before we get started it's helpful to understand the basic components of a valid 
PFA document. Every PFA document has an input spec, output spec, and an action. 
The reason for the specifications on the input and output is so that the machine 
executing your PFA can safely and reliably excute. PFA relies on the 
<a href="http://avro.apache.org/" target="_blank">Avro serialization format</a> 
for its type system, whichsupports your typical primitive data types like
boolean, integer, and string, but also, arrays, maps, and 
<a href="http://dmg.org/pfa/docs/avro_types/" target="_blank">more</a>. You can 
define these types in R by using any of the functions like `avro_*`.

```{r avro-types}
avro_double

avro_enum(list("one", "two"))
```

After specifying inputs and outputs, you must specify an action. With `aurelius` 
an action can be specified from an R expression. A simple example is `2+2`. This 
can be expanded to manipulate the input and/or derivatives of it. Here is a 
simple analytics spec that takes a double and adds 10 to it.

```{r trivial-model}
pfa_document(input = avro_double, 
             output = avro_double, 
             action = expression(input + 10))
```

You'll notice that the resulting output is a list. PFA is based on JSON, so it's 
natural for the document, when it's in R, to be a list that can always later be 
converted to JSON for export.

## Converting a Tree Model

Let's step through a more realistic example that predicts the species in the iris 
dataset, based on the other variables using a random forest model. First, format 
the columns to meet Avro guidelines: `[A-Za-z_][A-Za-z0-9_]*`. Having periods 
in the variable names (e.g. `Sepal.Length`) is not encouraged.

```{r format-cols}
# make sure column names dont have periods in them
iris2 <- iris
colnames(iris2) <- gsub('\\.', '_', colnames(iris2)) 
```

Creating the PFA is very easy. Simply fit your model, then call `pfa()`.

```{r fit-model}
# fit a random forest model
rf_model <- randomForest(Species ~ ., data=iris2) 
 
# convert that model to PFA
# pred_type='prob' means that the output is the 
# probability of all classes. If you want 
# the majority vote, then specify pred_type='response'
# check all options by running ?pfa.randomForest
rf_model_as_pfa <- pfa(rf_model, pred_type='prob') 
```

If you have the <a href="https://cran.r-project.org/web/packages/rPython/index.html" target="_blank">rPython</a> package
installed and <a href="https://github.com/opendatagroup/hadrian/wiki/Installation#case-4-you-want-to-install-titus-in-python" target="_blank">Titus</a> (PFA for Python), then you can create an engine in R 
and see how the PFA behaves. In this case it predicts the same as the `predict()` 
function in R. This way we can test that the behavior of the PFA is what we expect.

```{r check-model}
# you can check the predictions from this model
# by first converting to a pfa_engine
pfa_model <- pfa_engine(rf_model_as_pfa) 
 
# confirm that the pfa engine produces the 
# same prediction as the "predict" method in R
test_dat <- iris2[73,1:4]
round(pfa_model$action(as.list(test_dat)), 6)
round(unclass(predict(object=rf_model, newdata=test_dat, type='prob')), 6) 
```

Once you're happy with your model, just export to a flat file using `write_pfa()`.
This format can be read and used by any PFA interpreter. With R you can read the 
file back in using `read_pfa()`.

```{r read-write-pfa}
# you can export your model for use in other systems
write_pfa(rf_model_as_pfa, file = '~/my_rf_model.pfa')

my_model <- read_pfa(file("~/my_rf_model.pfa"))

# see that all PFA components are there
names(my_model)
```

## Final Thoughts 

PFA is a measurable step in open, portable, and reprodcible analytics. I hope to 
see the standard more widely adopted and tools developed to help others implement it. 
One final thing: Check out the 
<a href="https://github.com/StevenMMortimer/hadrian/tree/feature/add-r-package-structure/aurelius/tests/testthat" target="_blank">unit tests</a>. 
I've recently added quite a bit of unit test coverage and these tests are an 
excellent source of examples because they cover most all cases of utilizing the 
package functions.
